Supporting information: Copulas and their potential for ecology

### 2019-09-22 SupportingInformation.Rmd
replace copula abbreviations with names in appendix S13

We examined using simulations whether copula structure influences spatial Taylor's law, as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ replicate 
simulations of (temporally independent, in this example) 
spatio-temporal population dynamics measured in 
$n$ locations over $N$ time steps. We used $N=50$.
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of populations in 
each sampling location for each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *N-copula population array* 
for this simulation exercise;
copula structure of dependences between populations in different locations
was normal (N) for this array.

We then generated an series of arrays that were the same dimensions ($N \times n \times 1000$)
as the N-copula population array, and that were statistically similar to 
the N-copula population array in all ways except the new
arrays were constructed to have Clayton (respectively, Frank, survival Clayton) copula structure for 
the dependences between 
populations in different sampling locations within a simulation. We 
refer to the new arrays as our
*C-copula*, *F-copula*, and *SC-copula population arrays*, respectively. The C-copula array was
constructed as follows; the others were constructed analogously. First, we constructed an
$n$-dimensional Clayton copula with parameter chosen so that the 
Spearman correlations between each pair of the $n$ components were the same as those
of the normal distribution constructed above. This was done using the `rho` and `iRho` functions 
in the `copula` package in R. Because the Clayton copula is a one-paremeter family, this 
step relies on the fact that all pairs of locations were 
equally correlated with each other in the original multivariate normal distribution
constructed above. We then randomized each simulation in 
the N-copula population array in a special way 
(see Appendix \ref{sect:copsurrognd}) so that the resulting 
surrogate/randomized dataset had the copula structure
of the Clayton copula just constructed. The randomization procedure worked 
by permuting time series of values in the N-copula array, separately for each 
sampling location and simulation. So population marginal distrbutions for each location
and simulation were *exactly* the same as those in the N-copula population array. 
Permutations were done in such a way
that resulting pairwise Spearman correlations were the same as the C-copula 
constructed above, and hence were the same as the N-copula array, 
except for sampling variation. Thus, to summarize using mathematical notation,
if $m^{(\text{N})}$ is our N-copula population array and $m^{(\text{C})}$ is our 
C-copula population array, then,
for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
$\{m^{(\text{C})}_{tij}: t=1,\ldots,N\}$ are exactly the same (when considered 
as unordered sets); 
and for any given $i_1$, $i_2$ and $j$, the Spearman correlations $\cor_t(m^{(\text{C})}_{ti_1j},m^{(\text{C})}_{ti_2j})$
and $\cor_t(m^{(\text{N})}_{ti_1j},m^{(\text{N})}_{ti_2j})$ are the same, to within 
sampling variation. The whole exercise was repeated preserving Kendall instead of 
Spearman correlations.

@@

Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations in our N-copula population array), $p$-values for the linearty
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_t)$ versus $\log(m_t)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
relationship when copula structure was not normal.

>>>>

We examined using simulations whether copula structure influences spatial Taylor's law, as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ replicate 
simulations of (temporally independent, in this example) 
spatio-temporal population dynamics measured in 
$n$ locations over $N$ time steps. We used $N=50$.
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of populations in 
each sampling location for each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *normal-copula population array* 
for this simulation exercise;
copula structure of dependences between populations in different locations
was normal for this array.

We then generated a series of arrays that were the same dimensions ($N \times n \times 1000$)
as the normal-copula population array, and that were statistically similar to 
the normal-copula population array in all ways except the new
arrays were constructed to have Clayton (respectively, Frank, survival Clayton) copula structure for 
the dependences between 
populations in different sampling locations within a simulation. We 
refer to the new arrays as our
*Clayton-copula*, *Frank-copula*, and *Survival Clayton-copula population arrays*, respectively. The Clayton-copula array was
constructed as follows; the others were constructed analogously. First, we constructed an
$n$-dimensional Clayton copula with parameter chosen so that the 
Spearman correlations between each pair of the $n$ components were the same as those
of the normal distribution constructed above. This was done using the `rho` and `iRho` functions 
in the `copula` package in R. Because the Clayton copula is a one-paremeter family, this 
step relies on the fact that all pairs of locations were 
equally correlated with each other in the original multivariate normal distribution
constructed above. We then randomized each simulation in 
the normal-copula population array in a special way 
(see Appendix \ref{sect:copsurrognd}) so that the resulting 
surrogate/randomized dataset had the copula structure
of the Clayton copula just constructed. The randomization procedure worked 
by permuting time series of values in the normal-copula array, separately for each 
sampling location and simulation. So population marginal distrbutions for each location
and simulation were *exactly* the same as those in the normal-copula population array. 
Permutations were done in such a way
that resulting pairwise Spearman correlations were the same as the Clayton copula 
constructed above, and hence were the same as the normal-copula array, 
except for sampling variation. Thus, to summarize using mathematical notation,
if $m^{(\text{N})}$ is our normal-copula population array and $m^{(\text{C})}$ is our 
Clayton-copula population array, then,
for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
$\{m^{(\text{C})}_{tij}: t=1,\ldots,N\}$ are exactly the same (when considered 
as unordered sets); 
and for any given $i_1$, $i_2$ and $j$, the Spearman correlations $\cor_t(m^{(\text{C})}_{ti_1j},m^{(\text{C})}_{ti_2j})$
and $\cor_t(m^{(\text{N})}_{ti_1j},m^{(\text{N})}_{ti_2j})$ are the same, to within 
sampling variation. The whole exercise was repeated preserving Kendall instead of 
Spearman correlations.

@@

Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations in our normal-copula population array), $p$-values for the linearty
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_t)$ versus $\log(m_t)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
relationship when copula structure was not normal.
